Supplementary material for\\ A Benchmark of Named Entity Recognition Approaches in Historical Documents\\ Application to 19$^{th}$ Century French Directories

### 2022-01-20 src-latex/parts/30-dataset.tex
Fix bad section reference in sec. 3.

\subsection{A selection of Paris trade directories from 1798 to 1854}

The directories are available from different libraries in Paris and have been digitised independently in various levels of quality. 
They cover a large period of time and originate from several publishers.
Therefore, their contents, indexes, layouts, methods of printing, etc. may vary a lot from one directory to another (see Figure \ref{fig:directories}).
We want our groundtruth dataset to be representative of the diversity of directories available on the period.

@@

Directories contain lists of people with various information attached.
For instance the directory published by \emph{Didot} in 1854 contains three redundant lists of people sorted by name, by activity and by street name.
An typical example entry from this directory is "\textit{Batton (D.-A.) \scalerel*{\includegraphics{./images/LH.png}}{H}, professeur au Conservatoire de musique et de déclamation, Saint-Georges, 47.}".
It begins with the person's name and surname, here inside parentheses.
The glyph denotes that this person was awarded the \textit{Légion d'Honneur}.
Then comes a descriptions of the person's activity, here his profession (professor at the Conservatory of music and declamation), but it can also be a social status.
Such descriptions range from a single word to paragraphs describing the occupation in full detail.
The street name and number where the person lives or carries out their activities ends the entry.

These are the structured data that we want to extract, deduplicate  and structure to build a spatio-temporal database.
Except for some potentially wordy activity descriptions, they correspond to named entities. However, while most entries contain the same types of named entities, their order and the way they are written vary from one directory/index to another. To provide examples of each entry structure, pages from each type have thus been annotated. %\footnote{In this work, the street name index entries have not been included as they are very different from the others. The strategy to be adopted to process them without loss of quality for the NER model will be the subject of future work.}.

@@

The simplified data extraction pipeline depicted in~\cref{fig.pipeline} processes the documents presented
in~\cref{sec:dataset}. First, the page layout extraction and entry segmentation are performed with a
semi-automated system and checked by a human. The resulting images, representing 8765 entries, are the inputs of a customizable two-stage pipeline.

%1. Entrée/Sortie de l'OCR
\textbf{OCR stage.} The first stage of the pipeline aims at extracting the raw text from the images. An OCR system runs
on the thumbnails of each segmented entry to extract its text. An entry might span over multiple text lines but is
always a single block. Thus, the most adapted mode is chosen when the OCR system allows for the detection mode (e.g. the
\emph{block} mode for \emph{Tesseract}). The glyphs used in this dataset might be unknown to the OCR. Some of them, like
\scalerel*{\includegraphics{./images/LH.png}}{H}, do not even have a unicode codepoint. Text normalisation is thus required
on the OCR outputs and consists in projecting Unicode characters onto the latin-1 \emph{charset} (whenever it makes
sense), homogenise extra symbols (star and hand symbols) and alleviate some ambiguities (e.g. \emph{no} and \emph{n°}).

@@

%2. Entrée/Sortie
\textbf{NER Stage.} The second stage of the pipeline aims an extracting the named entities from a text with a NER
system. This text originates from the OCR outputs in a real-word scenario but might be the groundtruth text in order to
evaluate the NER performance independently. There are 5 type of entity to detect~(see.~\cref{tab.entities}).
The NER system has to classify non-overlapping parts of the text into one of these entity (or none of them). A second
Q.A. (namely NER Q.A.) is performed at the end of this stage between the groundtruth and the NER output.

Note that the dataset used to assess the NER stage is a subset of the entries. Indeed, we need to ensure that the datasets
contain the same entries whichever the OCR used in the previous stage. Entries where the OCR produced an empty string
and those for which no entity could be projected from the groundtruth have to be ignored. We filter the set of entries by
keeping only the entries that are always valid at the end of the stage 1. At the cost of losing 5\% of the references
entries, we thus have 8341 valid entries at the input of the stage 2.

@@

The evaluation process is illustrated on~\cref{fig.eval-ocr-ner}. The OCR and the groundtruth texts are first aligned to
evaluate the OCR accuracy. As there are 11 mismatches over 56 aligned characters, the CER is thus 24\%. This alignment
is then used to back-project the groundtruth tags to build the tagged NER-\emph{target}. Finally, the NER system runs on the
OCR text; its output is compared to the NER groundtruth. There is only 2 over 3 tags matching in the prediction (precision),
while only 2 over 4 entities are matched in the reference (recall). It follows an overall f-score of 0.4.
%
A side effect of tag projection is that it can produce degenerated tags, i.e. empty or containing only spaces.
In such cases, the corresponding entries must be removed from the dataset,
and only the intersection of the sets of valid entries must be used to compare NER systems with multiple OCR predictions or references.

>>>>

This section presents the historical sources that we selected and their contents. It also details the construction of the groundtruth dataset leveraged in our experiments and the metrics used to evaluate OCR and NER results.

\subsection{A selection of Paris trade directories from 1798 to 1854}
\label{sec:corpus}

The directories are available from different libraries in Paris and have been digitised independently in various levels of quality. 
They cover a large period of time and originate from several publishers.
Therefore, their contents, indexes, layouts, methods of printing, etc. may vary a lot from one directory to another (see Figure \ref{fig:directories}).
We want our groundtruth dataset to be representative of the diversity of directories available in the period.

@@

Directories contain lists of people with various information attached.
For instance, the directory published by \emph{Didot} in 1854 contains three redundant lists of people sorted by name, by activity, and by street name.
An typical example entry from this directory is "\textit{Batton (D.-A.) \scalerel*{\includegraphics{./images/LH.png}}{H}, professeur au Conservatoire de musique et de déclamation, Saint-Georges, 47.}".
It begins with the person's name and surname, here inside parentheses.
The glyph denotes that this person was awarded the \textit{Légion d'Honneur}.
Then comes a description of the person's activity, here his profession (professor at the Conservatory of music and declamation), but it can also be a social status.
Such descriptions range from a single word to paragraphs describing the occupation in full detail.
The street name and number where the person lives or carries out their activities ends the entry.

These are the pieces of information that we want to extract, deduplicate  and structure to build a spatio-temporal database.
Except for some potentially wordy activity descriptions, they correspond to named entities. However, while most entries contain the same types of named entities, their order and the way they are written vary from one directory/index to another. To provide examples of each entry structure, pages from each type have thus been annotated. %\footnote{In this work, the street name index entries have not been included as they are very different from the others. The strategy to be adopted to process them without loss of quality for the NER model will be the subject of future work.}.

@@

The simplified data extraction pipeline depicted in~\cref{fig.pipeline} processes the documents presented
in~\cref{sec:corpus}. First, the page layout extraction and entry segmentation are performed with a
semi-automated system and checked by a human. The resulting images, representing 8765 entries, are the inputs of a customizable two-stage pipeline.

%1. Entrée/Sortie de l'OCR
\textbf{OCR stage.} The first stage of the pipeline aims at extracting the raw text from the images. An OCR system runs
on the thumbnails of each segmented entry to extract its text. An entry might span over multiple text lines but is
always a single block. Thus, the most adapted mode is chosen when the OCR system allows for the detection mode (e.g. the
\emph{block} mode for \emph{Tesseract}). The glyphs used in this dataset might be unknown to the OCR. Some of them, like
\scalerel*{\includegraphics{./images/LH.png}}{H}, do not even have a Unicode codepoint. Text normalisation is thus required
on the OCR outputs and consists in projecting Unicode characters onto the latin-1 \emph{charset} (whenever it makes
sense), homogenise extra symbols (star and hand symbols), and alleviate some ambiguities (e.g. \emph{no} and \emph{n°}).

@@

%2. Entrée/Sortie
\textbf{NER Stage.} The second stage of the pipeline aims at extracting the named entities from a text with a NER
system. This text originates from the OCR outputs in a real-word scenario but might be the groundtruth text in order to
evaluate the NER performance independently. There are 5 types of entity to detect~(see.~\cref{tab.entities}).
The NER system has to classify non-overlapping parts of the text into one of these entities (or none of them). A second
Q.A. (namely, NER Q.A.) is performed at the end of this stage between the groundtruth and the NER output.

Note that the dataset used to assess the NER stage is a subset of the entries. Indeed, we need to ensure that the datasets
contain the same entries whichever the OCR used in the previous stage. Entries where the OCR produced an empty string
and those for which no entity could be projected from the groundtruth have to be ignored. We filter the set of entries by
keeping only the entries that are always valid at the end of stage 1. At the cost of losing 5\% of the references
entries, we thus have 8341 valid entries at the input of stage 2.

@@

The evaluation process is illustrated on~\cref{fig.eval-ocr-ner}. The OCR and the groundtruth texts are first aligned to
evaluate the OCR accuracy. As there are 11 mismatches over 56 aligned characters, the CER is thus 24\%. This alignment
is then used to back-project the groundtruth tags to build the tagged NER-\emph{target}. Finally, the NER system runs on the
OCR text; its output is compared to the NER groundtruth. There is only 2 over 3 tags matching in the prediction (precision),
while only 2 over 4 entities are matched in the reference (recall). It follows an overall f-score of 0.4.
%
A side effect of tag projection is that it can produce degenerated tags, i.e., empty or containing only spaces.
In such cases, the corresponding entries must be removed from the dataset,
and only the intersection of the sets of valid entries must be used to compare NER systems with multiple OCR predictions or references.
